---
title: ".slate[Communicating Economics]"
subtitle: ".reg[.slate[Data Visualisation in R:] Pre-Workshop Instructions]"
author: ".slate[Zaeen de Souza <br> Instructor: Surbhi Kesar]"
date: <br> .slate[2-05-2021]
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: clear

```{R, setup, include = F}
# devtools::install_github("dill/emoGG")
library(pacman)
p_load(
  broom, tidyverse,
  ggplot2, ggthemes, ggforce, ggridges,
  latex2exp, viridis, extrafont, gridExtra, kable,
  kableExtra, snakecase, janitor,
  data.table, dplyr,
  lubridate, knitr,
  estimatr, here, magrittr, tabulizer, rJava, dplyr)

library(readxl)
link <- read_excel("link.xlsx")


# Defining colors
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Dark slate grey: #314f4f
# Knitr options
opts_chunk$set(
  fig.align = "center",
  comment = "#",
  fig.height = 10,
  fig.width = 8,
  warning = F,
  message = F)
```

### Data

All the datasets that we will be using can be accessed [.hi[here]](https://drive.google.com/drive/u/1/folders/1lmL0JB0GYnSFtOD8FSC_cLWUkU9gBMju). Please download these beforehand, if you'd like to note down and run the code, as we go through the slides.



### .slate[Software]

For this workshop, you need to have both .purple.mono[`R`] and .purple.mono[`RStudio`] installed on your computer. You can download .purple.mono[`R`] from [.hi[here]](https://cloud.r-project.org/), and .purple.mono[`RStudio`] from [.hi[here]](https://www.rstudio.com/products/rstudio/download/), in case you don't have it already.  Some of the packages we will be using **require** a more recent version, so you might have to update it. (.mono.purple[ver.4.0]) and above will be ideal.

### .slate[Packages]

The following code can be used to install (or update) the packages we will be using for this workshop. The code can be copied using .pink.mono[`ctrl+c`], and pasted, using .pink.mono[`ctrl+v`], into a new .purple.mono[`R`] script file, and run using .pink.mono[`ctrl+a`], which selects all the code, and .pink.mono[`ctrl+enter`], which runs the selected code. Please do this before the workshop, in case you'd like to copy down and run the code as we go through the slides. Some basic details on the packages are available on the last page, along with links to some exmaples/explanations of the packages.

```{r, eval=F, echo=T, tidy=T}
# Use this to update R from within RStudio
install.packages("installr")
library(installr)
# This last command, will open up a download prompt; choose yes/no accordingly.
updateR()

# This is a package manager - it helps to load/install packages from the web.
# install it, if you don't have it already.
install.packages("pacman")
# Load it
library(pacman)

# load/and or install the packages we need for this. p_load will automatically install packages, if missing, and load, if already installed. Note: some of these will give you messages that look like warnings - its OK, as long as you have a new-ish version of R.
p_load(
tidyverse,
dplyr,
ggplot2,
ggthemes,
ggmap,
ggalluvial,
grid,
gridExtra,
gridtext,
extrafont,
extrafontdb,
Hmisc,
ggalt,
ggtext,
cowplot,
ggforce,
sf)
```

### .slate[General Reading]

If you're not fully comfortable with syntax, you could do a quick review [.hi[here]](https://rstudio-pubs-static.s3.amazonaws.com/307258_a6eb369630204f83b0e2860f36be20ec.html). It covers all the syntax required for the workshop, and as well a little more. I think it should take around 45-60 minutes to go through (if you do the exercises, as well). There is some very useful .mono.purple[`dplyr`] syntax, [.hi[here]](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html).

### For Stata Users

In addition to the above material, if needed, you can also look at some simple .purple.mono[`Stata`] to .purple.mono[`R`] translations [.hi[here]](https://github.com/EconometricsBySimulation/RStata/wiki/Dictionary:-Stata-to-R) and [.hi[here]](https://github.com/rstudio/cheatsheets/raw/master/stata2r.pdf).

.footer[Communicating Economics 2021]
---
class: clear

### Package Details

You can find the .mono.purple[`readme`] files for all the packages we will use, in this table below.

```{r, echo=F, eval=T}
kable(link) %>%
    kable_styling(font_size = 8)
```
.footer[Communicating Economics 2021]
